---
title: "Modelling"
author: "Xilai Fu"
date: "12/19/2020"
output: html_document
---

```{r}
library(nlme)
library(ggplot2)
library(nlme)
```

```{r}
load(file = "update_new_On_linear.Rda")
```

Our objective is to model the patients in linear group with linear regresson model and multilevel random effects model to best represent each the age group.


# Linear regression model (retrospective)
In this part, I will test the effect of Age group,Sex and their corresponding interaction terms with time.
Models are compared by anova test
```{r fig.width= 10,fig.height= 4.38,warning=FALSE}
## change reference categories of ajusted age and Sex
update_new_On_linear$ajusted_age <-factor(update_new_On_linear$ajusted_age,levels = c("c","a","b"))

levels(update_new_On_linear$Sex) <- c("M","F")  

model1 <- lm(Scores ~ Time.to.death, data = update_new_On_linear)
summary(model1)

model2 <- lm(Scores ~ Time.to.death+ajusted_age, data = update_new_On_linear)
summary(model2)

model3 <- lm(Scores ~ Time.to.death+ ajusted_age*Time.to.death+Sex*Time.to.death, data = update_new_On_linear)
summary(model3)

# test the effect of ajusted_age
anova(model1,model2,test = "Chisq")

# test the effect of interactions of time*sex and time*agegroup
anova(model2,model3,test = "Chisq")
```

# Multilevel random effects model (retrospective)
We are going to add a random intercept on ID to see if the model will be improved. multilevel random effects Models will be compared by likelihood ratio test.
```{r fig.width= 10,fig.height= 4.38}
mixed.lmer_1 <- lme(Scores ~ Time.to.death+ajusted_age, data = update_new_On_linear,method = 'ML',random=~1|ID)
summary(mixed.lmer_1)

anova(mixed.lmer_1,model2) # compared the model of mixed.lmer_1 and mode20 with likelihood ratio test

mixed.lmer_2 <- lme(Scores ~ Time.to.death+ajusted_age +Time.to.death:ajusted_age, data = update_new_On_linear,method = 'ML',random=~1|ID)
summary(mixed.lmer_2)

anova(mixed.lmer_1,mixed.lmer_2) # test the effect of Time*ajusted_age

mixed.lmer_2 <- lme(Scores ~ Time.to.death + Sex*Time.to.death, data = update_new_On_linear,method = 'ML',random=~1|ID)
summary(mixed.lmer_2) # test the effect of sex and its interaction term with time

mixed.lmer_3 <- lme(Scores ~ Time.to.death + ajusted_age +Time.to.death:ajusted_age, data = update_new_On_linear,method = 'ML',random=~1|ID)
summary(mixed.lmer_3)
anova(mixed.lmer_1,mixed.lmer_3) # test the effect of ajusted_age*time
```

# Model diagnostics
We are going to check the residual plots of the optimal linear model and multilevel random effect model.
```{r fig.width= 10,fig.height= 4.38}
## mixed.lmer_3
par(mfrow = c(1,2))

plot(mixed.lmer_3$fitted[,2],residuals(mixed.lmer_3),xlim=c(0,60),ylim = c(-15,15),frame = FALSE,xlab = "Fitted values",ylab = "Residuals")
abline(h = 0,col = "red")

qqnorm(resid(mixed.lmer_3), pch = 1, frame = FALSE,main = "")
qqline(resid(mixed.lmer_3), col = "red", lwd = 2)



## Model1.3
par(mfrow = c(1,2))

plot(model3$fitted,residuals(model3),xlim=c(0,60),ylim = c(-30,30),frame = FALSE,xlab = "Fitted values",ylab = "Residuals")
abline(h = 0,col = "red")

qqnorm(resid(model3), pch = 1, frame = FALSE,main = "",ylim = c(-30,30))
qqline(resid(model3), col = "red", lwd = 2)
```

# Secondary modeling
## Since the two age groups <50 and 50 - 64 are close in slope in the final model, we decided to have them assigned the same interaction term in a new model.
```{r}
## combine the categorical levels <50 and 50 - 64 for interaction term.
update_new_On_linear$age_indicator <- sapply(update_new_On_linear$ajusted_age,function(x){
  if(x %in% c("a","b")){
    return(1)
  }else{
    return(0)
  }
})

mixed.lmer <- lme(Scores ~ Time.to.death + ajusted_age +Time.to.death:age_indicator, data = update_new_On_linear,method = 'ML',random=~1|ID)

summary(mixed.lmer)
update_new_On_linear$age_indicator <- as.factor(update_new_On_linear$age_indicator)
levels(update_new_On_linear$age_indicator)

anova(mixed.lmer_1,mixed.lmer) ## likelihood ratio test comparing the new interaction term age_indicator.time 
```



# Modelling from time of death
Visualization of multilevel random effects model with its fixed effects.
```{r}
## Extract the time range for each age group
range(update_new_On_linear[update_new_On_linear$ajusted_age == "a","Time.to.death"])
range(update_new_On_linear[update_new_On_linear$ajusted_age == "b","Time.to.death"])
range(update_new_On_linear[update_new_On_linear$ajusted_age == "c","Time.to.death"])

## Extract the coefficients of the fixed effect
intercept <- coef(summary(mixed.lmer))[1,1]
Time <- coef(summary(mixed.lmer))[2,1] 
Agegroup1 <- coef(summary(mixed.lmer))[3,1] 
Agegroup2 <- coef(summary(mixed.lmer))[4,1] 
interact <- coef(summary(mixed.lmer))[5,1]

## Construct the equations with the associated coefficients obtained from our final model for each age group

# >= 65
equation1=function(x){Time*x+intercept}
# < 50
equation2=function(x){(Time+interact)*x+intercept+Agegroup1}
# 50 - 64
equation3=function(x){(Time+interact)*x+intercept+Agegroup2}

p1 <- ggplot(update_new_On_linear, aes(y = Scores, x = Time.to.death)) +
  stat_function(fun = equation2,
                geom = "line",
                xlim = c(-13, 0),
                aes(color = "a")) +
  stat_function(fun = equation3,
                geom = "line",
                xlim = c(-11, 0),
                aes(color = "b")) +
  stat_function(fun = equation1,
                geom = "line",
                xlim = c(-11, 0),
                aes(color = "c")) +
  
  
  scale_colour_manual(
    "Age at Diagnosis (PD, n = 29)",
    values = c("a" = "#FC4E07", "b" = "#E7B800", "c" = "#00AFBB"),
    labels = c("a" = "< 50", "b" = "50 - 65", "c" = greatequal65)
  ) +
  
  geom_segment(
    y = 0,
    yend = 60,
    x = Inf,
    xend = Inf,
    lwd = 0.5,
    colour = "grey30",
    lineend = "square"
  ) +
  geom_segment(
    y = -Inf,
    yend = -Inf,
    x = -13,
    xend = 0,
    lwd = 0.5,
    colour = "grey30",
    lineend = "square"
  ) +
  labs(x = "\nTime prior to death(years)", y = "Scores(UPDRSIII)\n") +
  scale_x_continuous(limits = c(-16, 0), breaks = seq(-16, 0, 1)) +
  scale_y_continuous(
    limits = c(0, 65),
    breaks = seq(0, 65, 5),
    position = "right"
  ) + theme(
    axis.line = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.title = element_text(size = 12.8, color = "black"),
    legend.text = element_text(size = 12.8),
    axis.text.x = element_text(size = 12.8, face = "plain"),
    axis.text.y = element_text(size = 12.8, face = "plain"),
    axis.title = element_text(size = 13.8, face = "plain"),
    axis.ticks.length = unit(.18, "cm"),
    legend.justification = c(0, 1),
    legend.position = c(0.018, 0.998),
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  coord_cartesian(ylim = c(0, 60), xlim = c(-13, 0))

p1
```
