---
title: "UPDRS data cleaning"
author: "Xilai Fu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    number_sections: true
    css: style.css
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true 
      smooth_scroll: true
---
```{r}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```



# Package
```{r setup, include=FALSE}
# Packages -----
library(stringr)
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lattice)
library(magrittr)
library(miscTools)
library(tidyverse)
library(plyr)
library(splines)
library(MASS)
library(lspline)
library(ggrepel)
```



# Functions
```{r setup, include=FALSE}
####Extract num from string
numextract<-function(string){
  num1<-0
  if(grepl("months",string)){
    num1<- as.numeric(str_extract(string,"\\-*\\d+\\.*\\d*"))
    num1<-num1/12
  }else{
    num1<-as.numeric(str_extract(string,"\\-*\\d+\\.*\\d*"))
  }
  return(num1)
}

# find row name 
index_fun <- function(dat, target, name){
  position <- (dat[, name] == target)
  rownames(dat)[position]
}


# ggplot text(they ended up being useless for my work)
lm_eqn1 <- function(df){
  m <- lm(Scores ~ Distance, df);
  eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(R)^2~"="~r2, 
                   list(a = format(unname(coef(m)[1]), digits = 2),
                        b = format(unname(coef(m)[2]), digits = 2),
                        r2 = format(summary(m)$r.squared, digits = 3)))
  as.character(as.expression(eq));
}

lm_eqn2 <- function(df){
  m <- lm(Scores ~ Distance+I(Distance^2), df);
  eq <- substitute(italic(y) == a + b %.% italic(x)+c %.% italic(x^2)*","~~italic(R)^2~"="~r2, 
                   list(a = format(unname(coef(m)[1]), digits = 2),
                        b = format(unname(coef(m)[2]), digits = 2),
                        c = format(unname(coef(m)[3]), digits = 2),
                        r2 = format(summary(m)$r.squared, digits = 3)))
  
  
  as.character(as.expression(eq));
}

lm_eqn3 <- function(df){
  m <- lm(Scores ~ Distance+I(Distance^2)+I(Distance^3), df);
  eq <- substitute(italic(y) == a + b %.% italic(x)+c %.% italic(x^2)+d %.% italic(x^3)*","~~italic(R)^2~"="~r2, 
                   list(a = format(unname(coef(m)[1]), digits = 2),
                        b = format(unname(coef(m)[2]), digits = 2),
                        c = format(unname(coef(m)[3]), digits = 2),
                        d = format(unname(coef(m)[4]), digits = 2),
                        r2 = format(summary(m)$r.squared, digits = 3)))
  
  
  as.character(as.expression(eq));
}

```


* Read data
```{r }
data = read.csv(file="UPDRS1updated.csv",header=TRUE,fileEncoding="latin1")
data <-data[data$Dx=="PD",]
```

# Data Cleaning
## Primary data manipulation/cleaning
```{r eval=FALSE, include=FALSE}
## create the ID and delete MMSE and MOCA
a <- seq(13, 141, 4)
b <- seq(14, 142, 4)
ID = paste(data$Number,data$Initial,sep="") 
data1 <- subset(data, select = -c(a,b))
columns = subset(data1,select=-c(1,2))
data2 = cbind(ID = ID,columns)

## Convert all units("Year"","month"") into "Year".(Temporary) for wide formate data
data2$Time.to.first.UPDRS.III<-as.character(data2$Time.to.first.UPDRS.III)
data2$Time.to.first.UPDRS.III<-sapply(data2$Time.to.first.UPDRS.III,function(x){as.numeric(numextract(x))})
data2$Time.to.first.UPDRS.III<-as.numeric(data2$Time.to.first.UPDRS.III)
data2$Time.to.first.UPDRS.III<-round(data2$Time.to.first.UPDRS.III,2)

## Spilit the data into two parts and change them into required format
s<-seq(11,75,2)
t<-seq(10,74,2)
sub_str <-subset(data2,select =-s)
sub_num <-subset(data2,select =c(1,s))

## Change the wide format into the long format for sub1
substr_long <- gather(sub_str, visitTimes, Distance, Year:Year.32, factor_key = TRUE)
p<-seq(11,42,1)
substr_long <- substr_long[order(substr_long$ID), ]


## Change the wide format into the long format for sub2
subnum_long <- gather(sub_num,UPDRS.III, Scores, UPDRS.III:UPDRS.III.32, factor_key = TRUE)
subnum_long <- subnum_long[order(subnum_long$ID), ]


## Combine sub1_long and sub2_long
extra <-subset(subnum_long,select=c(2,3))
totalData <-cbind(substr_long,extra)
totalData$visitTimes <- as.character(totalData$visitTimes)


## Convert the visitTimes into numeric
visitTimes<-unique(totalData$visitTimes)
totalData$visitTimes<-sapply(totalData$visitTimes, function(x){which(visitTimes %in% x)})

## Excluding rows with empty scores/Distance(filter 1,9 patients were dropped because of no observations)
updatedData<-totalData[!is.na(totalData$Scores), ]
updatedData<-updatedData[updatedData$Scores!="", ]
updatedData<-updatedData[!is.na(updatedData$Distance), ]

### Excluding Year.Dx(mark)
updatDat1 <-subset(updatedData,select = -c(5))

updatDat1$Time.to.first.UPDRS.III[updatDat1$ID=="127LF"] <-"3.5"
updatDat1$Time.to.first.UPDRS.III<-as.character(updatDat1$Time.to.first.UPDRS.III)
#length(unique(updatDat1$ID))


## (On/off manipulation)
for(j in 1:length(updatDat1$ID)){
  if(grepl("ON/OFF",updatDat1$Score[j]) ==TRUE){
    updatDat1$text[j]<-"ON/OFF"
    }else if(grepl("OFF meds ON DBS",updatDat1$Score[j],"OFF meds ON DBS")==TRUE){
    updatDat1$text[j]<-"(OFF meds ON DBS)"
    }else if(grepl("(OFF meds ON optimised DBS)",updatDat1$Score[j])==TRUE){
    updatDat1$text[j]<-"(OFF meds ON optimised DBS)"
    }else if(grepl("OFF",updatDat1$Score[j])==TRUE){
    updatDat1$text[j]<-"OFF"  
    }else{
    updatDat1$text[j]<-"ON"
  }
}

## Extract numeric part of the scores
updatDat1$Scores<-sapply(updatDat1$Scores,function(x){as.numeric(numextract(x))})

## unique manipulation
updatDat1[updatDat1$ID=="127LF",]$Time.to.first.UPDRS.III<-3.5
```

```{r}
## Age cut
data2$Year.Dx<-as.integer(str_extract(data2$Year.Dx,"\\-*\\d+\\.*\\d*"))
data2$Age.at.Dx<-as.integer(data2$Age.at.Dx)
data2$Age.at.Dx1<-cut(data2$Age.at.Dx,c(25,35,45,55,65,75,85,95),right = FALSE)
data2$gap<-cut(data2$Time.to.first.UPDRS.III,c(0,0.0001,1,5,10,15,20,25,30,35,39),right = FALSE)
```


## Secondary data cleaning
### data split by diagnosis and filter stage1
Since we are only interested in **ON medication** score modeling for patients, we decided to drop the scores that are not **ON medication**
```{r time, eval=FALSE, include=FALSE}
## data type deletion and type conversion
updatDat1 <-
  updatDat1 %>% dplyr::select(-c(visitTimes, UPDRS.III, DOB..DD.MM.YYYY.))
updatDat1$Scores<-as.numeric(updatDat1$Scores)
updatDat1$Time.to.first.UPDRS.III<-as.numeric(updatDat1$Time.to.first.UPDRS.III)
updatDat1$Distance<-as.integer(updatDat1$Distance)

## ON/OFF manipulation
newRow<-updatDat1[14,]
row1<-updatDat1[1:14,]
row2<-updatDat1[-c(1:14),]
updatDat1<-rbind(row1,newRow,row2)
updatDat1$Scores[14]<-"24.5"
updatDat1$Scores[15]<-"46"
updatDat1$text[14]<-"ON"
updatDat1$text[15]<-"OFF"
updatDat1$Scores<-as.numeric(updatDat1$Scores)
length(unique(updatDat1$ID))
## drop all the scores with non On medication and retain only on medication scores(filter stage 2)
## 17MR is dropped because it does not have on medication socres
updatDat_on <-updatDat1[updatDat1$text =="ON",]
updatDat_on$text <-NULL

## data split by diagnosis((PD and LBD),(PSP,MSA)) and save it to updatDat1
updatDat_1<-updatDat_on
dat_PL<-updatDat_on[updatDat_on$Dx %in% c("PD","LBD"),]
dat_PM_cleaned <-updatDat_on[updatDat_on$Dx %in% c("PSP","MSA"),]
```

### subset(PD and LBD) manipulation
```{r time, eval=FALSE, include=FALSE}
dat_PL_cleaned <-
  dat_PL %>%  aggregate(by = list(dat_PL$ID, dat_PL$Sex, dat_PL$Dx, dat_PL$Distance),drop=TRUE, mean)
dat_PL_cleaned<-subset(dat_PL_cleaned,select=-c(4,5,7,10))
dat_PL_cleaned<-dat_PL_cleaned[order(dat_PL_cleaned$Group.1,dat_PL_cleaned$Distance),]
names(dat_PL_cleaned)[1]<-"ID"
names(dat_PL_cleaned)[2]<-"Sex"
names(dat_PL_cleaned)[3]<-"Dx"


## Important Variables derivation
### Post diagnosis distance in time 
dat_PL_cleaned$Distance1 <-dat_PL_cleaned$Time.to.first.UPDRS.III + dat_PL_cleaned$Distance
### Age at visit
dat_PL_cleaned$Age.at.visit<-dat_PL_cleaned$Age.at.Dx+dat_PL_cleaned$Distance1
### Year at visit
dat_PL_cleaned$Year.at.visit<-as.numeric(dat_PL_cleaned$Year.of.Birth)+dat_PL_cleaned$Age.at.visit
### Time t0 death
dat_PL_cleaned$Time.to.death<-dat_PL_cleaned$Year.at.visit-dat_PL_cleaned$Year.death
### reset visitTimes
names<-unique(dat_PL_cleaned$ID)
dat_PL_cleaned$visitTimes<-0
for(i in names){
  for(j in 1:length(dat_PL_cleaned[dat_PL_cleaned$ID==i,"ID"])){
    dat_PL_cleaned[dat_PL_cleaned$ID==i,]$visitTimes[j]<-j
  }
}
```

#### filter the patients(PD and LBD,stage2)
In filter stage2, we disregard the patients with illogical time to death and those whose clinical observations were not available in their last two years of post diagnosis life
```{r}
## clean for PD and LBD
### 7HS and 137MH both have postive time from death, which is medically illogical(dropped)
#dat_PL_cleaned<-dat_PL_cleaned[(dat_PL_cleaned$ID!="7HS")&(dat_PL_cleaned$ID!="137MH"),]
dat_PL_cleaned<-dat_PL_cleaned[dat_PL_cleaned$ID!="137MH",]

## Pnames is the set of patients with positive Time.to.death
Pnames<-unique(dat_PL_cleaned[dat_PL_cleaned$Time.to.death>0,]$ID)

## Adjustment for positive time.to.death
options(scipen = 999)
for(id in Pnames){
  dat_PL_cleaned[dat_PL_cleaned$ID==id,]$Time.to.death<-dat_PL_cleaned[dat_PL_cleaned$ID==id,]$Time.to.death-max(dat_PL_cleaned[dat_PL_cleaned$ID==id,]$Time.to.death)
}

## pl_un is a set of patients 
pl_un<-c()
ID <- unique(dat_PL_cleaned$ID)
for(id in ID){
  sub<-dat_PL_cleaned[dat_PL_cleaned$ID==id,]
  if(max(sub$Time.to.death)<=0&max(sub$Time.to.death)>= -2){
    pl_un<-c(pl_un,id)
  }
}

dat_PL_cleaned_2<-dat_PL_cleaned[dat_PL_cleaned$ID%in%pl_un,]

length(pl_un)
```


#### filter the patients(PD and LBD,stage3)
filter the patients who had at least 5 post diagnosis observations
```{r time, eval=FALSE, include=FALSE}
## filter the patients who had at least 5 post diagnosis observations
uni_Id<-unique(dat_PL_cleaned_2$ID)
IDRS<-c()
for(i in uni_Id){
  if(length(unique(dat_PL_cleaned_2[dat_PL_cleaned_2$ID==i,]$Distance))>=5){
    IDRS<-c(IDRS,i)
    
  }}
length(IDRS)
dat_PL_cleaned_3<-dat_PL_cleaned_2[dat_PL_cleaned_2$ID%in%IDRS,]
#save(update_new,file = "update_new.Rda")
#load(file = "update_new.Rda")
length(unique(dat_PL_cleaned_3$ID))

```

### subset(MSA and PSP) manipulation
#### filter the patients(PSP and MSA,stage2)
filter the patients who had observations in their last two years of life
```{r}
# data manipulation for MSA&PSP
dat_PM_cleaned<-dat_PM
dat_PM_cleaned$Distance1<-as.numeric(dat_PM_cleaned$Distance)+as.numeric(dat_PM_cleaned$Time.to.first.UPDRS.III)
length(unique(dat_PM_cleaned$ID))
# distance ajustment
for(i in 1:length(dat_PM_cleaned$ID) ){
  if(i>=4){
   C1<-dat_PM_cleaned$ID[i]==dat_PM_cleaned$ID[i-1]&dat_PM_cleaned$Distance1[i]==dat_PM_cleaned$Distance1[i-1]
   C2<-dat_PM_cleaned$ID[i-1]==dat_PM_cleaned$ID[i-2]&dat_PM_cleaned$Distance1[i-1]==dat_PM_cleaned$Distance1[i-2]
   C3<-dat_PM_cleaned$ID[i-2]==dat_PM_cleaned$ID[i-3]&dat_PM_cleaned$Distance1[i-2]==dat_PM_cleaned$Distance1[i-3]
   if(C1&C2&C3){
     dat_PM_cleaned$Distance1[i-2]<-as.numeric(dat_PM_cleaned$Distance1[i-2])+0.25
     dat_PM_cleaned$Distance1[i-1]<-as.numeric(dat_PM_cleaned$Distance1[i-1])+0.5
     dat_PM_cleaned$Distance1[i]<-as.numeric(dat_PM_cleaned$Distance1[i])+0.75
     }else if(C2&C3){
     dat_PM_cleaned$Distance1[i-2]<-as.numeric(dat_PM_cleaned$Distance1[i-2])+0.33333333
     dat_PM_cleaned$Distance1[i-1]<-as.numeric(dat_PM_cleaned$Distance1[i-1])+0.66666666
     }else if(C3){
      
     dat_PM_cleaned$Distance1[i-2]<-as.numeric(dat_PM_cleaned$Distance1[i-2])+1/2
     }
 }
   
  }

dat_PM_cleaned$Age.at.visit<-dat_PM_cleaned$Age.at.Dx+dat_PM_cleaned$Distance1
dat_PM_cleaned$Year.at.visit<-dat_PM_cleaned$Year.of.Birth+dat_PM_cleaned$Age.at.visit
dat_PM_cleaned$Time.to.death<-dat_PM_cleaned$Year.at.visit-dat_PM_cleaned$Year.death

## cleaned total MP dataset
dat_PM_cleaned[dat_PM_cleaned$ID=="21ES",]$Time.to.death<-dat_PM_cleaned[dat_PM_cleaned$ID=="21ES",]$Time.to.death-max(dat_PM_cleaned[dat_PM_cleaned$ID=="21ES",]$Time.to.death)
dat_PM_cleaned[dat_PM_cleaned$ID=="42BM",]$Time.to.death<-dat_PM_cleaned[dat_PM_cleaned$ID=="42BM",]$Time.to.death-	max(dat_PM_cleaned[dat_PM_cleaned$ID=="42BM",]$Time.to.death)

#save(datMP,file = "datMP.Rda")
#load(file = "datMP.Rda")

## pm_un is a set of patients who had observations in their last two years 
pm_un<-c()
ID <- unique(dat_PM_cleaned$ID)
for(id in ID){
  sub<-dat_PM_cleaned[dat_PM_cleaned$ID==id,]
  if(max(sub$Time.to.death)<=0&max(sub$Time.to.death)>= -2){
    pm_un<-c(pm_un,id)
  }
}

### reset visitTimes
names<-unique(dat_PM_cleaned$ID)
dat_PM_cleaned$visitTimes<-0
for(i in names){
  for(j in 1:length(dat_PM_cleaned[dat_PM_cleaned$ID==i,"ID"])){
    dat_PM_cleaned[dat_PM_cleaned$ID==i,]$visitTimes[j]<-j
  }
}


dat_PM_cleaned_2<-dat_PM_cleaned[dat_PM_cleaned$ID%in%pm_un,]
```

#### filter the patients(PSP and MSA,stage3)
filter the patients who had at least 5 observations
```{r}
MPname<-c()
for(i in unique(dat_PM_cleaned_2$ID)){
  if(length(unique(dat_PM_cleaned_2[dat_PM_cleaned_2$ID==i,]$Distance1))>=5){
    MPname<-c(MPname,i)
    
  }}

dat_PM_cleaned_3<-dat_PM_cleaned_2[dat_PM_cleaned_2$ID%in%MPname,]
```

# save important data sets
```{r}
dat_PL_cleaned_2$group<-"PL"
dat_PM_cleaned_2$group<-"PM"

dat_PL_cleaned_3$group<-"PL"
dat_PM_cleaned_3$group<-"PM"

full_data<-updatDat1
analysis_data1<-rbind(dat_PL_cleaned_2,dat_PM_cleaned_2)
analysis_data2<-rbind(dat_PL_cleaned_3,dat_PM_cleaned_3)


save(full_data,file = "new data/full_data.Rda")
save(analysis_data1,file = "new data/analysis_data1.Rda")
save(analysis_data2,file = "new data/analysis_data2.Rda")

dat_PD_cleaned<-dat_PL_cleaned[dat_PL_cleaned$Dx=="PD",]
save(dat_PD_cleaned,file = "new data/dat_PD_cleaned.Rda")
```



