---
title: "UPDRS Project"
author: "Xilai Fu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    number_sections: true
    css: style.css
    code_folding: hide
    toc: true
    toc_float:
      collapsed: true 
      smooth_scroll: true
---
```{r}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r setup, include=FALSE}
# Packages -----
library(stringr)
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lattice)
library(magrittr)
library(miscTools)
library(tidyverse)
library(plyr)
library(splines)
library(MASS)
library(lspline)
library(ggrepel)
library(stargazer)
library(boot)
library(table1)
library(kableExtra)
library(arsenal)
library(magrittr)
```


```{r}
load(file = "full_data.Rda")
load(file = "analysis_data1.Rda")
load(file = "analysis_data2.Rda")
```

# Table1.1 Descriptive statistics of full data and analysis I analysis II data.
## Descriptive statistics of full data
```{r}
#full_data
#analysis_data1
#analysis_data2
full_data_table<-full_data[,c(1,3,4,6,7)]
full_data_wide<-unique(full_data_table)
full_data_wide$`Number of Visits`<- sapply(full_data_wide$ID,function(x){nrow(full_data[full_data$ID==x,])})

## Sex
full_data_wide$Sex<-sapply(full_data_wide$Sex,function(x){if(x == "M"){x<-"Male"}else{x<-"Female"}})

## Diagnosis
full_data_wide$Dx<-sapply(full_data_wide$Dx,function(x){if(x == "PD"){x<-"Parkindon's Disease"}else if(x == "LBD"){x<-"Lewy body dementia"}else if(x == "MSA"){x<-"Multiple system atrophy"}else{x<-"Progressive supranuclear palsy"}})

## Age cut and additon of corrsponding column
full_data_wide$`Age at Diagnosis`<-cut(full_data_wide$Age.at.Dx,c(27,50,65,90),right = TRUE)
age_class <-levels(full_data_wide$`Age at Diagnosis`)
full_data_wide$`Age at Diagnosis`<-sapply(full_data_wide$`Age at Diagnosis`,function(x){if(x == age_class[1]){x <- "<=50"}else if(x == age_class[2]){x<-"50-<=65"}else{x<-">65"}})


## Time.from..diagnosis.to.first.UPDRSIII cut and addition of corresponding column
full_data_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-cut(full_data_wide$Time.to.first.UPDRS.III,c(-1,0.0000001,1,5,10,15,100),right = TRUE)
gap_class <-levels(full_data_wide$`Time from diagnosis to first UPDRS-III assessment(years)`)

full_data_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-sapply(full_data_wide$`Time from diagnosis to first UPDRS-III assessment(years)`,function(x){if(x == gap_class[1]){x <- "0"}else if(x == gap_class[2]){x<-"0-<=1"}else if(x == gap_class[3]){x<-"1-<=5"}else if(x == gap_class[4]){x<-"5-<=10"}else if(x == gap_class[5]){x<-"10-<=15"}else{x<-">15"}})


## Number of visits and addition of corresponding column
full_data_wide$`Number of Visits_1`<-cut(full_data_wide$`Number of Visits`,c(0,5,10,15,100),right = FALSE)

visit_class<-levels(full_data_wide$`Number of Visits_1`)
full_data_wide$`Number of Visits_1`<-sapply(full_data_wide$`Number of Visits_1`,function(x){if(x == visit_class[1]){x <- "<5"}else if(x == visit_class[2]){x<-"5-<10"}else if(x == visit_class[3]){x<-"10-<15"}else{x<-">=15"}})


## reorder levels of categorical levels
### Sex
full_data_wide$Sex<-factor(full_data_wide$Sex,levels = c("Male","Female"))
### Dignosis
full_data_wide$Dx<-factor(full_data_wide$Dx,levels = c("Parkindon's Disease","Lewy body dementia","Multiple system atrophy","Progressive supranuclear palsy"))
### Age at Diagnosis
full_data_wide$`Age at Diagnosis`<- factor(full_data_wide$`Age at Diagnosis`,levels = c("<=50","50-<=65",">65"))
### Time from diagnosis to first UPDRSIII
full_data_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-factor(full_data_wide$`Time from diagnosis to first UPDRS-III assessment(years)`,levels = c("0","0-<=1","1-<=5","5-<=10","10-<=15",">15"))
### Number of visits
full_data_wide$`Number of Visits_1`<- factor(full_data_wide$`Number of Visits_1`,levels = c("<5","5-<10","10-<15",">=15"))

### label and units
label(full_data_wide$Dx)<-"Diagnosis"
```

## Descriptive statistics of analysis data I
```{r}
analysis_data1_table<-analysis_data1[,c(1,2,3,5,7)]
analysis_data1_wide<-unique(analysis_data1_table)

analysis_data1_wide$`Number of Visits`<- sapply(analysis_data1_wide$ID,function(x){nrow(full_data[full_data$ID==x,])})

## Sex
analysis_data1_wide$Sex<-sapply(analysis_data1_wide$Sex,function(x){if(x == "M"){x<-"Male"}else{x<-"Female"}})

## Diagnosis
analysis_data1_wide$Dx<-sapply(analysis_data1_wide$Dx,function(x){if(x == "PD"){x<-"Parkindon's Disease"}else if(x == "LBD"){x<-"Lewy body dementia"}else if(x == "MSA"){x<-"Multiple system atrophy"}else{x<-"Progressive supranuclear palsy"}})

## Age cut and additon of corrsponding column
analysis_data1_wide$`Age at Diagnosis`<-cut(analysis_data1_wide$Age.at.Dx,c(27,50,65,90),right = TRUE)
age_class <-levels(analysis_data1_wide$`Age at Diagnosis`)
analysis_data1_wide$`Age at Diagnosis`<-sapply(analysis_data1_wide$`Age at Diagnosis`,function(x){if(x == age_class[1]){x <- "<=50"}else if(x == age_class[2]){x<-"50-<=65"}else{x<-">65"}})


## Time.from..diagnosis.to.first.UPDRSIII cut and addition of corresponding column
analysis_data1_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-cut(analysis_data1_wide$Time.to.first.UPDRS.III,c(-1,0.0000001,1,5,10,15,100),right = TRUE)
gap_class <-levels(analysis_data1_wide$`Time from diagnosis to first UPDRS-III assessment(years)`)

analysis_data1_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-sapply(analysis_data1_wide$`Time from diagnosis to first UPDRS-III assessment(years)`,function(x){if(x == gap_class[1]){x <- "0"}else if(x == gap_class[2]){x<-"0-<=1"}else if(x == gap_class[3]){x<-"1-<=5"}else if(x == gap_class[4]){x<-"5-<=10"}else if(x == gap_class[5]){x<-"10-<=15"}else{x<-">15"}})


## Number of visits and addition of corresponding column
analysis_data1_wide$`Number of Visits_1`<-cut(analysis_data1_wide$`Number of Visits`,c(0,5,10,15,100),right = FALSE)

visit_class<-levels(analysis_data1_wide$`Number of Visits_1`)
analysis_data1_wide$`Number of Visits_1`<-sapply(analysis_data1_wide$`Number of Visits_1`,function(x){if(x == visit_class[1]){x <- "<5"}else if(x == visit_class[2]){x<-"5-<10"}else if(x == visit_class[3]){x<-"10-<15"}else{x<-">=15"}})


## reorder levels of categorical levels
### Sex
analysis_data1_wide$Sex<-factor(analysis_data1_wide$Sex,levels = c("Male","Female"))
### Dignosis
analysis_data1_wide$Dx<-factor(analysis_data1_wide$Dx,levels = c("Parkindon's Disease","Lewy body dementia","Multiple system atrophy","Progressive supranuclear palsy"))
### Age at Diagnosis
analysis_data1_wide$`Age at Diagnosis`<- factor(analysis_data1_wide$`Age at Diagnosis`,levels = c("<=50","50-<=65",">65"))
### Time from diagnosis to first UPDRSIII
analysis_data1_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-factor(analysis_data1_wide$`Time from diagnosis to first UPDRS-III assessment(years)`,levels = c("0","0-<=1","1-<=5","5-<=10","10-<=15",">15"))
### Number of visits
analysis_data1_wide$`Number of Visits_1`<- factor(analysis_data1_wide$`Number of Visits_1`,levels = c("<5","5-<10","10-<15",">=15"))

### label and units
label(analysis_data1_wide$Dx)<-"Diagnosis"
```

## Descriptive statistics of analysis data II
```{r}
analysis_data2_table<-analysis_data2[,c(1,2,3,5,7)]
analysis_data2_wide<-unique(analysis_data2_table)

analysis_data2_wide$`Number of Visits`<- sapply(analysis_data2_wide$ID,function(x){nrow(full_data[full_data$ID==x,])})

## Sex
analysis_data2_wide$Sex<-sapply(analysis_data2_wide$Sex,function(x){if(x == "M"){x<-"Male"}else{x<-"Female"}})

## Diagnosis
analysis_data2_wide$Dx<-sapply(analysis_data2_wide$Dx,function(x){if(x == "PD"){x<-"Parkindon's Disease"}else if(x == "LBD"){x<-"Lewy body dementia"}else if(x == "MSA"){x<-"Multiple system atrophy"}else{x<-"Progressive supranuclear palsy"}})

## Age cut and additon of corrsponding column
analysis_data2_wide$`Age at Diagnosis`<-cut(analysis_data2_wide$Age.at.Dx,c(27,50,65,90),right = TRUE)
age_class <-levels(analysis_data2_wide$`Age at Diagnosis`)
analysis_data2_wide$`Age at Diagnosis`<-sapply(analysis_data2_wide$`Age at Diagnosis`,function(x){if(x == age_class[1]){x <- "<=50"}else if(x == age_class[2]){x<-"50-<=65"}else{x<-">65"}})


## Time.from..diagnosis.to.first.UPDRSIII cut and addition of corresponding column
analysis_data2_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-cut(analysis_data2_wide$Time.to.first.UPDRS.III,c(-1,0.0000001,1,5,10,15,100),right = TRUE)
gap_class <-levels(analysis_data2_wide$`Time from diagnosis to first UPDRS-III assessment(years)`)

analysis_data2_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-sapply(analysis_data2_wide$`Time from diagnosis to first UPDRS-III assessment(years)`,function(x){if(x == gap_class[1]){x <- "0"}else if(x == gap_class[2]){x<-"0-<=1"}else if(x == gap_class[3]){x<-"1-<=5"}else if(x == gap_class[4]){x<-"5-<=10"}else if(x == gap_class[5]){x<-"10-<=15"}else{x<-">15"}})


## Number of visits and addition of corresponding column
analysis_data2_wide$`Number of Visits_1`<-cut(analysis_data2_wide$`Number of Visits`,c(0,5,10,15,100),right = FALSE)

visit_class<-levels(analysis_data2_wide$`Number of Visits_1`)
analysis_data2_wide$`Number of Visits_1`<-sapply(analysis_data2_wide$`Number of Visits_1`,function(x){if(x == visit_class[1]){x <- "<5"}else if(x == visit_class[2]){x<-"5-<10"}else if(x == visit_class[3]){x<-"10-<15"}else{x<-">=15"}})


## reorder levels of categorical levels
### Sex
analysis_data2_wide$Sex<-factor(analysis_data2_wide$Sex,levels = c("Male","Female"))
### Dignosis
analysis_data2_wide$Dx<-factor(analysis_data2_wide$Dx,levels = c("Parkindon's Disease","Lewy body dementia","Multiple system atrophy","Progressive supranuclear palsy"))
### Age at Diagnosis
analysis_data2_wide$`Age at Diagnosis`<- factor(analysis_data2_wide$`Age at Diagnosis`,levels = c("<=50","50-<=65",">65"))
### Time from diagnosis to first UPDRSIII
analysis_data2_wide$`Time from diagnosis to first UPDRS-III assessment(years)`<-factor(analysis_data2_wide$`Time from diagnosis to first UPDRS-III assessment(years)`,levels = c("0","0-<=1","1-<=5","5-<=10","10-<=15",">15"))
### Number of visits
analysis_data2_wide$`Number of Visits_1`<- factor(analysis_data2_wide$`Number of Visits_1`,levels = c("<5","5-<10","10-<15",">=15"))

### label and units
label(analysis_data2_wide$Dx)<-"Diagnosis"
```


## full data table
```{r}
table1(~ Sex + `Time from diagnosis to first UPDRS-III assessment(years)`+`Age at Diagnosis` + `Number of Visits_1`,caption = "Table1.1 Descriptive statistics of full data",data=full_data_wide)
full_data_wide<-full_data_wide[!(full_data_wide$ID %in% c("17MR","137MH")),]
```

## analysis data I table
```{r}
table1(~ Sex + Dx + `Time from diagnosis to first UPDRS-III assessment(years)`+`Age at Diagnosis` + `Number of Visits_1`,caption = "Table1.1 Descriptive statistics of analysis data1",data=analysis_data1_wide)
```

## analysis data II table
```{r}
table1(~ Sex + Dx + `Time from diagnosis to first UPDRS-III assessment(years)`+`Age at Diagnosis` + `Number of Visits_1`,caption = "Table1.1 Descriptive statistics of analysis data2",data=analysis_data2_wide)
```


## Maximum clinical table
```{r}
## get the mode of a vector
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

full_data_wide <- full_data_wide[full_data_wide$Dx=="Parkindon's Disease",]
## get the maximum number of annual visits for patients
full_names<-unique(full_data_wide$ID)
max_visits <- sapply(full_names, function(id) {
  mod <- getmode(full_data[full_data$ID == id, ]$Distance)
  max_num <-
    length(full_data[(full_data$ID == id) &
                       (full_data$Distance == mod), ]$ID)
  
  })

## combine ID and max visits
dat_max<-data.frame("ID"=full_names,"Max visits" = max_visits)
dat_Dx<-full_data_wide[,c(4)]
merge<-cbind(dat_max,dat_Dx)

max_visits<-as.factor(max_visits)
table1(~ dat_Dx|max_visits,caption = "Table1.2 Maximum clinical visits in a year in the full data",data=merge)
```

## Time interval from last visit to date of death in the full data
```{r}
full_data$last_visit_death
full_data_single<-subset(full_data,select = c(1:8))
full_data_single$max_dis <-
  sapply(full_data_single$ID, function(id) {
    max(full_data_single[full_data_single$ID == id, ]$Distance)
  })

full_data_max_dis<-unique(subset(full_data_single,select = -c(8)))

full_data_max_dis$death_gap <-
  full_data_max_dis$ID %>% sapply(function(id) {
    full_data_max_dis[full_data_max_dis$ID == id, ]$Year.death -          full_data_max_dis[full_data_max_dis$ID ==
                                                                                              id, ]$Year.of.Birth - full_data_max_dis[full_data_max_dis$ID == id, ]$Time.to.first.UPDRS.III -
      full_data_max_dis[full_data_max_dis$ID == id, ]$max_dis - full_data_max_dis[full_data_max_dis$ID == id, ]$Age.at.Dx
    
  })

## derive the year of last visit to death
full_data_max_dis$death_gap<-full_data_max_dis$death_gap %>% sapply(function(x){
  if((x>=-1)&(x<0)){
    return(0)
  }else{
    return(x)
  }
})

## categorize the year of last visit to death
full_data_max_dis$death_gap_range<-full_data_max_dis$death_gap %>% sapply(function(x){
  if(x<0){
    return("Positive/illogical")
  }else if((x>=0)&(x<=1)){
    return("Within the last year")    
  }else if((x>1)&(x<=2)){
    return("Within the second last year")
  }else{
    return("More than 2 years")
  }
})

full_data_max_dis$death_gap_range<-factor(full_data_max_dis$death_gap_range,c("Within the last year","Within the second last year","More than 2 years","Positive/illogical"))

table1(~ death_gap_range,caption = "Time interval from last visit to date of death in the full data",data=full_data_max_dis)
```

## Distribution of age from diagnosis in the full data
```{r}
dis1 <-as.numeric(full_data_max_dis$max_dis)+full_data_max_dis$Time.to.first.UPDRS.III
full_data_max_dis$age_from_death<-dis1 %>% sapply(function(x){
  if(x>=5){
    return(">=5 years")
  }else{
    return("<5 years")
  }
})

table1(~ age_from_death,caption = "Distribution of age from diagnosis in the full data",data=full_data_max_dis)
```







































































































# Statistics of modelling (PD)
## linear trajectory
```{r}
linear_name<-c("112GR","24EH","30MD","4FZ","50RH","59HP","67AL","86WG","117GT","121FF","146JVK","147RR","172VS","190KJ","191JH","217WH","101HP","103LC","122MP","140MC","56HT","88LO","129RE","166SS","189EC","213DR","222LB","102JB")

linear<-analysis_data2_PD[analysis_data2_PD$ID%in%linear_name,]

## unique row of each patient
linear_un <-unique(linear[,c(1,2,4,5)])

## unique names of PD patients in the linear group 
PD_linear_na<-unique(linear$ID)

## derive standard deviation
linear_un$sd <- sapply(PD_linear_na, function(id) {
  return(sd(linear[linear$ID == id, "Scores"]))
})

## derive R2
linear_un$R2 <- sapply(PD_linear_na, function(id) {
  f1<-lm(Scores ~ Distance1, data=linear[linear$ID== id,])
  r2<-summary(f1)$r.squared
  return(r2)
})

## derive slope
linear_un$slope <- sapply(PD_linear_na, function(id) {
  f1<-lm(Scores ~ Distance1, data=linear[linear$ID== id,])
  slope<-summary(f1)$coefficient[2,1]
  return(slope)
})

# Discriptive statistics of linear group
kable(summary(linear_un[,c("Age.at.Dx","sd","R2","slope")]))
non_linear[16]

non_linear[39]

non_linear[42]


non_linear <- unique(analysis_data2[!(analysis_data2$ID %in% c(linear_name,stable_name)),]$ID)
r2<-sapply(non_linear, function(id) {
  f1<-lm(Scores ~ Distance1, data=analysis_data2[analysis_data2$ID== id,])
  r2<-summary(f1)$r.squared
  return(r2)
})
```



## low variance trajectory
Since the observed standard deviation of 201GT is very low, we decided to move it to low variance group temporarily.
```{r}
stable_name<-c("26SG","219PF","143MS","203MZ","201GT")
stable <- analysis_data2_PD[analysis_data2_PD$ID %in% stable_name,]

## unique row of each patient
stable_un <-unique(stable[,c(1,2,4,5)])

## unique names of PD patients in the linear group 
PD_stable_na<-unique(stable_un$ID)

## derive standard deviation
stable_un$sd <- sapply(PD_stable_na, function(id) {
  return(sd(stable[stable$ID == id, "Scores"]))
})

## summary statistics of standard deviation and Age.at.Dx in low variance group
summary(stable_un)

```


## one knot
```{r}
## one knot
### increase-decrease
increase_decrease_name<-c("144LS","52DL","156EF","110LG")
increase_decrease_knot<-c(16.333333,5.3,7.3,10)
increase_decrease_type<-rep("increase_decrease",4)

increase_decrease_age.dx <- sapply(increase_decrease_name, function(x) {
  return(analysis_data2_wide[analysis_data2_wide$ID == x, ]$Age.at.Dx)
})

dat_ID <-
  data.frame(
    "ID" = increase_decrease_name,
    "Age_at_Dx" = increase_decrease_age.dx,
    "knot_1" = increase_decrease_knot,
    "trajectory_type" = increase_decrease_type
  )

### flat-increase
flat_increase_name<-c("6GT","13MQ","179EH","186WD","205JW","16ER","93EP")
flat_increase_knot<-c(7,19,5,25,12.83333,3.333333,9)
flat_increase_type<-rep("flat_increase",7)
flat_increase_age.dx <- sapply(flat_increase_name, function(x) {
  return(analysis_data2_wide[analysis_data2_wide$ID == x, ]$Age.at.Dx)
})

dat_FI<-data.frame("ID" = flat_increase_name,"Age_at_Dx"=flat_increase_age.dx,"knot_1" = flat_increase_knot,"trajectory_type" =flat_increase_type)

### decrease-increase
decrease_increase_name<-c("66EL","72EO","82EF","136BM","84MT","131TB","171JW","105DG","194PK")
decrease_increase_knot<-c(2,4,4,21.5,12,9,8,7,3)
decrease_increase_type<-rep("decrease_increase",9)
decrease_increase_age.dx <- sapply(decrease_increase_name, function(x) {
  return(analysis_data2_wide[analysis_data2_wide$ID == x, ]$Age.at.Dx)
})

dat_DI <-
  data.frame(
    "ID" = decrease_increase_name,
    "Age_at_Dx" = decrease_increase_age.dx,
    "knot_1" = decrease_increase_knot,
    "trajectory_type" = decrease_increase_type
  )

## cobine the dat of piecewise(one knot) trajectories
dat_one<-rbind(dat_ID,dat_FI,dat_DI)
dat_one$ID<-factor(dat_one$ID,levels = levels(analysis_data2_PD$ID))

## derive standard deviation
dat_one$sd <-
  sapply(dat_one$ID, function(id) {
    
    sd <- sd(analysis_data2_PD[analysis_data2_PD$ID == id,"Scores"])
    return(sd)
    
    
  })

## derive R2
dat_one$R2 <-
  sapply(dat_one$ID, function(id) {
    knot<-dat_one[dat_one$ID==id,]$knot_1
    sub<- analysis_data2_PD[analysis_data2_PD$ID == id,]
    
    f1 <- lm(Scores ~ lspline(Distance1, c(knot), marginal = TRUE),data = sub)
    
    r2 <- summary(f1)$r.squared
    return(r2)
    
    
  })

## derive the slope of the first segament
dat_one$slope_1 <-
  sapply(dat_one$ID, function(id) {
    knot<-dat_one[dat_one$ID==id,]$knot_1
    sub<- analysis_data2_PD[analysis_data2_PD$ID == id,]
    
    f1 <- lm(Scores ~ lspline(Distance1, c(knot), marginal = TRUE),data = sub)

    slo1 <- summary(f1)$coefficients[2, 1]
   
    return(slo1)
    
    
  })

## derive the slope of the second segament
dat_one$slope_2 <-
  sapply(dat_one$ID, function(id) {
    knot<-dat_one[dat_one$ID==id,]$knot_1
    sub<- analysis_data2_PD[analysis_data2_PD$ID == id,]
    
    f1 <- lm(Scores ~ lspline(Distance1, c(knot), marginal = TRUE),data = sub)
    
    slo2 <- summary(f1)$coefficients[3, 1]
    
    return(slo2)
    
    
  })


dat_one$knot1_significance <- sapply(dat_one$ID, function(id) {
  f0 <-
    lm(Scores ~ Distance1, data = analysis_data2[analysis_data2$ID == id,])
  f1 <-
    lm(Scores ~ lspline(Distance1, c(dat_one[dat_one$ID == id,]$knot_1), marginal =
                          TRUE), data = analysis_data2[analysis_data2$ID == id,])
  
  return(anova(f0, f1)[6][2, 1])
})
  



## summary of Age.at.Dx,knot,sd,R2,slope1 and slope2 by trajectory type
stby(dat_one[,c(2:3,5:8)], dat_one$trajectory_type, descr, stats = "fivenum") %>%
  tb(order = 3)


### stable portion of flat_increase
FI <-dat_one[dat_one$trajectory_type=="flat_increase",]

FI_ID<- FI$ID

flat_portion<-sapply(FI_ID,function(id){
  knot<-FI[FI$ID==id,]$knot_1
  dat_point<-analysis_data2_PD[(analysis_data2_PD$ID==id)&(analysis_data2_PD$Distance1<=knot),]$Scores
  return(sd(dat_point))
})


## summary
summary(flat_portion)
```
### p_value graph 
```{r}
## pvalue assignment and cut for dat_one
dat_one$significance<- cut(dat_one$knot1_significance,c(0,0.05,0.12),right = FALSE)
dat_one$significance_text<-sapply(dat_one$significance,function(x){
  if(x == "[0,0.05)"){
    return("significant")
  }else{
    return("not significant")
  } 
})

## pvalue assignment and cut for dat_two
dat_two$significance<- cut(dat_two$knot12_significance,c(0,0.05,0.5),right = FALSE)
dat_two$significance_text<-sapply(dat_two$significance,function(x){
  if(x == "[0,0.05)"){
    return("significant")
  }else{
    return("not significant")
  } 
})

## R_squa
linear_un$R2_significance<- cut(linear_un$R2,c(0,0.7,1),right = FALSE)


## export data

write.csv(linear_un,file = "new data/linear.csv")
write.csv(dat_one,file = "new data/dat_one.csv")
write.csv(dat_two,file = "new data/dat_two.csv")
```


### one knot model comparison
```{r}
f0 <- lm(Scores ~ Distance1, data=analysis_data2[analysis_data2$ID ==dat_one$ID[7], ])

f1 <- lm(Scores ~ lspline(Distance1, c(dat_one$knot_1[7]), marginal=TRUE), data=analysis_data2[analysis_data2$ID ==dat_one$ID[7], ])


anova(f0,f1)[6][2,1]
```


## two knots
```{r}
# two knots
## increase_decrease_increase
increase_decrease_increase_name <- c("20MS","76HF","79FF","157DD","168RW","130RC","94AP")
increase_decrease_increase_knot1 <-c(10,23,3,12,10,12,7)
increase_decrease_increase_knot2 <-c(14,25,4,13,13,13,9)
increase_decrease_increase_type<-rep("increase_decrease_increase",7)
increase_decrease_increase_age.dx <-
  sapply(increase_decrease_increase_name, function(x) { 
    return(analysis_data2_wide[analysis_data2_wide$ID == x,]$Age.at.Dx)
  })


### subset_IDI 
dat_IDI <-
  data.frame(
    "ID" = increase_decrease_increase_name,
    "Age_at_Dx" = increase_decrease_increase_age.dx,
    "knot_1" = increase_decrease_increase_knot1,
    "knot_2" = increase_decrease_increase_knot2,
    "trajectory_type" = increase_decrease_increase_type
  )


## decrease_increase_decrease
decrease_increase_decrease_name <- c("22PH","120LW","175HL")
decrease_increase_decrease_knot1 <-c(6,17,3.75)
decrease_increase_decrease_knot2 <-c(14,20,6.75)
decrease_increase_decrease_type<-rep("decrease_increase_decrease",3)
decrease_increase_decrease_age.dx <-
  sapply(decrease_increase_decrease_name, function(x) { 
    return(analysis_data2_wide[analysis_data2_wide$ID == x,]$Age.at.Dx)
  })


### subset_DID 
dat_DID <-
  data.frame(
    "ID" = decrease_increase_decrease_name,
    "Age_at_Dx" = decrease_increase_decrease_age.dx,
    "knot_1" = decrease_increase_decrease_knot1,
    "knot_2" = decrease_increase_decrease_knot2,
    "trajectory_type" = decrease_increase_decrease_type
  )

dat_two<-rbind(dat_IDI,dat_DID)

dat_two$ID<-factor(dat_two$ID,levels = levels(analysis_data2_PD$ID))

dat_two$sd <-
  sapply(dat_two$ID, function(id) {
    sd <- sd(analysis_data2_PD[analysis_data2_PD$ID == id,"Scores"])
    
    return(sd)
    
    
  })

## derive R2
dat_two$R2 <-
  sapply(dat_two$ID, function(id) {
    knot_1 <- dat_two[dat_two$ID == id, ]$knot_1
    knot_2 <- dat_two[dat_two$ID == id, ]$knot_2
    sub <- analysis_data2_PD[analysis_data2_PD$ID ==id,]
    
    f1 <-
      lm(Scores ~ lspline(Distance1, c(knot_1,knot_2), marginal = TRUE), data = sub)
    
    r2 <- summary(f1)$r.squared
    return(r2)
    
    
  })

## derive the slope of the first segament
dat_two$slope_1 <-
  sapply(dat_two$ID, function(id) {
    knot_1 <- dat_two[dat_two$ID == id, ]$knot_1
    knot_2 <- dat_two[dat_two$ID == id, ]$knot_2
    sub <- analysis_data2_PD[analysis_data2_PD$ID ==id,]
    
    f1 <-
      lm(Scores ~ lspline(Distance1, c(knot_1,knot_2), marginal = TRUE), data = sub)

    slo1 <- summary(f1)$coefficients[2, 1]
   
    return(slo1)
    
    
  })


## derive the slope of the second segament
dat_two$slope_2 <-
  sapply(dat_two$ID, function(id) {
    knot_1 <- dat_two[dat_two$ID == id, ]$knot_1
    knot_2 <- dat_two[dat_two$ID == id, ]$knot_2
    sub <- analysis_data2_PD[analysis_data2_PD$ID ==id,]
    
    f1 <-
      lm(Scores ~ lspline(Distance1, c(knot_1,knot_2), marginal = TRUE), data = sub)

    slo2 <- summary(f1)$coefficients[3, 1]
   
    return(slo2)
    
    
  })



## derive the slope of the third segament
dat_two$slope_3 <-
  sapply(dat_two$ID, function(id) {
    knot_1 <- dat_two[dat_two$ID == id, ]$knot_1
    knot_2 <- dat_two[dat_two$ID == id, ]$knot_2
    sub <- analysis_data2_PD[analysis_data2_PD$ID ==id,]
    
    f1 <-
      lm(Scores ~ lspline(Distance1, c(knot_1,knot_2), marginal = TRUE), data = sub)
    
    slo3 <- summary(f1)$coefficients[4, 1]
    
    return(slo3)
    
    
  })

dat_two$knot12_significance <- sapply(dat_two$ID, function(id) {
  f0 <-
    lm(Scores ~ Distance1, data = analysis_data2[analysis_data2$ID == id,])
  
  f1 <-
    lm(Scores ~ lspline(Distance1, c(dat_two[dat_two$ID == id,]$knot_1,dat_two[dat_two$ID == id,]$knot_2), marginal =TRUE), data = analysis_data2[analysis_data2$ID == id,])
  
  return(anova(f0, f1)[6][2, 1])
})


## summary of Age.at.Dx,knot,sd,R2,slope1 and slope2 by trajectory type
stby(dat_two[,c(2:4,6:10)], dat_two$trajectory_type, descr, stats = "fivenum") %>%
  tb(order = 3)
```

## irregular/undetermined
```{r}
all_fitted_name <- c(linear_name,stable_name,increase_decrease_name,flat_increase_name,decrease_increase_name,increase_decrease_increase_name,decrease_increase_decrease_name)

## data of patients with irregular trajectories
dat_irre<- analysis_data2_PD[!(analysis_data2_PD$ID%in% all_modeled),]

irr_ID<-unique(dat_irre$ID)
irre_sd<-sapply(irr_ID, function(id) {
    sd <- sd(analysis_data2_PD[analysis_data2_PD$ID == id,"Scores"])
    
    return(sd)
    })

irre_age.dx <-
  sapply(irr_ID, function(x) { 
    return(analysis_data2_wide[analysis_data2_wide$ID == x,]$Age.at.Dx)
  })

dat_irre <-
  data.frame("ID" = irr_ID,
             "Age_at_Dx" = irre_age.dx,
             "trajectory_type" = "irregular")

summary(irre_age.dx)
summary(irre_sd)
```

## trajectory groups by sex 
```{r}
linear_un$trajectory_type<-"linear"
stable_un$trajectory_type<-"low variance"

## collective data of 0 knot
dat_0 <- rbind(linear_un[,c(1,2,4,8)],stable_un[,c(1,2,4,6)])

## collective data of knot(s)
dat_knot<-rbind(dat_one[,c(1,2,4)],dat_two[,c(1,2,5)])

## data of irregular trajectories
dat_irre


## add age.at.Dx for dat_knot and dat_irre
dat_age <- rbind(dat_knot,dat_irre)

dat_age$Sex <- sapply(dat_age$ID,function(id){
  return(analysis_data2_wide[analysis_data2_wide$ID==id,]$Sex)
})

## total data
colnames(dat_0)[3] <-"Age_at_Dx"
dat_total <-rbind(dat_0,dat_age)



## Age cut
dat_total$Age_cut<-cut(dat_total$Age_at_Dx,c(27,50,65,90),right = TRUE)
age_class <-levels(dat_total$Age_cut)

lesequa50<-stringi::stri_replace_all_fixed(
  c("<= 50"), 
  c("<="), 
  c("\u2264"), 
  vectorize_all = F
)

lesequa65<-stringi::stri_replace_all_fixed(
  c("50 -<= 65"), 
  c("<="), 
  c("\u2264"), 
  vectorize_all = F
)

dat_total$Age_category<-sapply(dat_total$Age_cut,function(x){if(x == age_class[1]){x <- lesequa50}else if(x == age_class[2]){x<-lesequa65}else{x<-"> 65"}})

dat_total$Sex<-as.character(dat_total$Sex)

## Sex adjustment
for(id in dat_total$ID) {
  if (dat_total[dat_total$ID == id, ]$Sex == "M") {
    dat_total[dat_total$ID == id, ]$Sex <- "Male"
  } else if (dat_total[dat_total$ID == id, ]$Sex == "F") {
     dat_total[dat_total$ID == id, ]$Sex <- "Female"
  }
}
dat_total$Sex<-as.factor(dat_total$Sex)

## adjust trajectory_type
dat_total$trajectory_type<- as.factor(dat_total$trajectory_type)

## adjust trajectory_type
dat_total$trajectory_type<-factor(dat_total$trajectory_type,levels = c("linear","low variance","increase_decrease","flat_increase","decrease_increase","increase_decrease_increase","decrease_increase_decrease","irregular"))

## adjust sex
dat_total$Sex<-factor(dat_total$Sex,levels=c("Male","Female"))

table1(~ trajectory_type|Sex,caption = "Table5.1 Trajectory groups by sex",data=dat_total)

```
## trajectory groups by age group
```{r}
table1(~ trajectory_type|Age_category,caption = "Table5.2 Trajectory groups by age group",data=dat_total)

```
## trajectory groups by age at diagnosis
```{r}
table1(~ Age_at_Dx|trajectory_type,caption = "Table5.2 Trajectory groups by age at diagnosis",data=dat_total)

stby(dat_total[,c(3,4)], dat_total$trajectory_type, descr, stats = "fivenum") %>%
  tb(order = 3)
```
